Effectful Semantics for Substructural SSA

### 2023-11-16 pldi24/paper.tex
Tap

Finally, we must give semantics for \rle{where}. To do so, we begin by giving a
interpreting a control-flow graph \(\lhaslb{\ms{L}}{L}{\ms{K}}\) as a morphism
from \(\dnt{\ms{L}}\) to \(\dnt{\ms{K}} + \dnt{\ms{L}}\), where the left-hand
side of the coproduct denotes branching to one of the target labels in
\(\ms{K}\) (the \textit{output}), while the right-hand side of the coproduct
denotes a recursive call to a label in \(\ms{L}\) (the \textit{recursive
input}). In particular, \rle{nil-br} simply treats a call to one of \(\ms{L}\)
as a call to a \textit{target label}, passing it through the left injection
unchanged. On the other hand \rle{cons-br} is a little bit more complicated:
first we execute \(\dnt{\lhaslb{\ms{L}}{L}{\ms{K}, \llhyp{\ell}{\Gamma}{A}}}\),
which will give us either an output \(\dnt{\ms{K}, \lhyp{\ell}{\Gamma}{A}}\), or
a recursive input \(\dnt{\ms{L}}\), the latter of which we can pass through
unchanged as a recursive input. In the output case, we can interpret this as
either a value of \(\dnt{\ms{K}}\) or \(\dnt{\Gamma} \otimes \dnt{A}\); the
former we can pass the value through unchanged as output; in the second case, we
can pass this in as input to \(\dnt{\haslb{\Gamma,
\thyp{x}{A}{q}}{t}{\ms{L}}}\), which will give us a value of \(\ms{L}\) we can
pass in as recursive input. Note that, even if \(t\) branches to an output label
in \(\ms{K}\), we still return a \textit{recursive input}, not an output; in
this case, the \textit{next} iteration will return it as an output. We can now
quite easily state the semantics of \rle{where}: we simply precompose the
semantics of the inner block \(\dnt{\haslb{\Gamma}{t}{\ms{L}}}\) with the
\textit{iterated} semantics of the associated control-flow graph,
\(\dnt{\lhaslb{\ms{L}}{L}{\ms{K}}}^\dagger\).

>>>>

Finally, we must give semantics for \rle{where}. To do so, we begin by giving a
interpreting a control-flow graph \(\lhaslb{\ms{L}}{L}{\ms{K}}\) as a morphism
from \(\dnt{\ms{L}}\) to \(\dnt{\ms{K}} + \dnt{\ms{L}}\), where the left-hand
side of the coproduct denotes branching to one of the target labels in
\(\ms{K}\) (the \textit{output}), while the right-hand side of the coproduct
denotes a recursive call to a label in \(\ms{L}\) (the \textit{recursive
input}). In particular, \rle{nil-br} simply treats a call to one of \(\ms{L}\)
as a call to a \textit{target label}, passing it through the left injection
unchanged. On the other hand \rle{cons-br} is a little bit more complicated:
first we execute \(\dnt{\lhaslb{\ms{L}}{L}{\ms{K}, \llhyp{\ell}{\Gamma}{A}}}\),
which will give us either an output \(\dnt{\ms{K}, \lhyp{\ell}{\Gamma}{A}}\), or
a recursive input \(\dnt{\ms{L}}\), the latter of which we can pass through
unchanged as a recursive input. In the output case, we can interpret this as
either a value of \(\dnt{\ms{K}}\) or \(\dnt{\Gamma} \otimes \dnt{A}\); the
former we can pass the value through unchanged as output; in the second case, we
can pass this in as input to \(\dnt{\haslb{\Gamma,
\thyp{x}{A}{q}}{t}{\ms{L}}}\), which will give us a value of \(\ms{L}\) we can
pass in as recursive input. Note that, even if \(t\) branches to an output label
in \(\ms{K}\), we still return a \textit{recursive input}, not an output; in
this case, the \textit{next} iteration will return it as an output. We can now
quite easily state the semantics of \rle{where}: we simply precompose the
semantics of the entry block \(\dnt{\haslb{\Gamma}{t}{\ms{L}}}\) with the
\textit{iterated} semantics of the associated control-flow graph,
\(\dnt{\lhaslb{\ms{L}}{L}{\ms{K}}}^\dagger\).
